# Data Sources

## NYPD Arrest Data 

*(Year to Date)*

*Source: https://data.cityofnewyork.us/Public-Safety/NYPD-Arrest-Data-Year-to-Date-/uip8-fykc*

**Data Provided By**: New York Police Department

**Total records**: 140,413

**Type of variables**: 

**Issues with the data**:


```{r}
df_arrest_ytd <- readr::read_csv("datasets/NYPD_Arrest_Data__Year_to_Date_.csv")
str(df_arrest_ytd)
```

## NYPD Hate Crimes

*Source: https://data.cityofnewyork.us/Public-Safety/NYPD-Hate-Crimes/bqiq-cu78*
**Description**: Dataset containing confirmed hate crime incidents in New York City. A hate crime is a traditional offense like murder, arson, or vandalism with an added element of bias. For the purposes of collecting statistics, the FBI has defined a hate crime as a “criminal offense against a person or property motivated in whole or in part by an offender’s bias against a race, religion, disability, sexual orientation, ethnicity, gender, or gender identity.”  

**Data Provided By**: New York Police Department

**Total records**: 728

**Issues with the data**: County names are historical.
```{r}
df_hate_crimes <- readr::read_csv("datasets/NYPD_Hate_Crimes.csv")
```

**Type of variables**: 
```{r}
str(df_hate_crimes)
```

## NYC Population Data

### New York City Population By Neighborhood

*Source: https://data.cityofnewyork.us/City-Government/New-York-City-Population-By-Neighborhood-Tabulatio/swpk-hqdp*

The data was collected from Census Bureaus' Decennial data dissemination (SF1).
Neighborhood Tabulation Areas (NTAs), are aggregations of census tracts that are subsets of New York City's 55 Public Use Microdata Areas (PUMAs). Primarily due to these constraints, NTA boundaries and their associated names may not definitively represent neighborhoods.
This report shows change in population from 2000 to 2010 for each NTA.
Compiled by the Population Division – New York City Department of City Planning.

Column description:
- Standard Code Census Bureau uses to process data. The code is unique to all geographical units.
   - 005-Bronx, 047-Brooklyn, 061-Manhattan, 081-Queens, 085-Staten Island."
- Each NTA within NYC's 5 boroughs is given a unique value. 
- Neighborhood Tabulation Areas, or NTAs, are aggregations of census tracts that are subsets of New York City's 55 Public Use Microdata Areas (PUMAs).

```{r}
df_population_yearly <- readr::read_csv("datasets/New_York_City_Population_By_Neighborhood_Tabulation_Areas.csv")
str(df_population_yearly)
```

### Projected population per age group

*Source: https://data.cityofnewyork.us/City-Government/Projected-Population-2010-2040-Total-By-Age-Groups/97pn-acdf*

```{r}
df_population_age <- readr::read_csv("datasets/Projected_Population_2010-2040_-_Total_By_Age_Groups.csv")
str(df_population_age)
```

### Initial Questions

 - This census data can help identify the areas with higher crimes based on percentage of its own Population

## NYC Shooting Data 

### NYPD Shooting Incident Data (Year To Date)

*Source: https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Year-To-Date-/5ucz-vwe8*

```{r}
df_shooting_ytd <- readr::read_csv("datasets/NYPD_Shooting_Incident_Data__Year_To_Date_.csv")
str(df_shooting_ytd)
```

 - This is a breakdown of every shooting incident that occurred in NYC during the current calendar year. (2020) This data is manually extracted every quarter and reviewed by the Office of Management Analysis and Planning before being posted on the NYPD website.
Each record represents a shooting incident in NYC and includes information about the event, the location and time of occurrence. 
 - In addition, information related to suspect and victim demographics is also included. For example, has information about the sex, Age, Race of the victim.

### NYPD Shooting Incident Historic Data

*Source: https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Historic-/833y-fsy8*

```{r}
df_shooting_old <- readr::read_csv("datasets/NYPD_Shooting_Incident_Data__Historic_.csv")
str(df_shooting_old)
```
 
 
 - This is a breakdown of every shooting incident that occurred in NYC going back to 2006 through the end of the previous calendar year.(2019) This data is manually extracted every quarter and reviewed by the Office of Management Analysis and Planning before being posted on the NYPD website. Each record represents a shooting incident in NYC and includes information about the event, the location and time of occurrence. In addition, information related to suspect and victim demographics is also included.
 
### Initial Questions

 - The information about the shootings in across locations and multiple demographics can help identify groups more susceptible.
 - Further, Historical data of the shootings can help identify dangerous  neighborhoods for different groups. Possible hot-spots to avoid.
 - Also, might help analyze the overall trend in shooting across NYC over the years/months/day-of-week/time-of-day.

## FBI Human Trafficking Data

*Source: https://crime-data-explorer.fr.cloud.gov/downloads-and-docs*

```{r}
df_fbi_ht <- readr::read_csv("datasets/HT_2013_2019.csv")
str(df_fbi_ht)
```

*Many other important datasets can be found on this website.*

## FBI Drug Arrests Data

This dataset contains monthly number of arrests for drug abuse violations reported by participating law enforcement agencies from 1995–2016. The arrests are by offense and broken down by age and sex or age and race.

*(Reported Number of Drug Arrests)*

*Source: https://crime-data-explorer.fr.cloud.gov/downloads-and-docs*

**Data Provided By**: FBI

**Total records**: 23

**Issues with the data**: None although there is no state-wise information.

```{r}
df_fbi_drug_arrests <- readr::read_csv("datasets/arrests_national_drug.csv")
```
**Type of variables**: 
```{r}
str(df_fbi_drug_arrests)
```

## NYC citizen Census Data

*Source: https://datausa.io/profile/geo/new-york-ny#economy*

```{r}
df_nyc_earnings<- readr::read_csv("https://datausa.io/api/data?measure=Household%20Income%20by%20Race,Household%20Income%20by%20Race%20Moe&Geography=16000US3651000:parents,16000US3651000,16000US3651000:similar")
str(df_nyc_earnings)
```

## NYC park crime data

*Source: https://www1.nyc.gov/site/nypd/stats/crime-statistics/park-crime-stats.page*
**Description**: This is a dataset containing crimes in the parks of New York City. This dataset is gleaned from quarterly reports from the NYPD. The data transformation is detailed in the cleaning file. 

**Data Provided By**: New York Police Department

**Total records**: 27696

**Issues with the data**: None

```{r}
df_parks<- readr::read_csv("datasets/parks/finalpark.csv")
```
**Type of variables**: 
```{r}
str(df_parks)
```