# Data Sources

## NYPD Arrest Data 

*(Year to Date)*

*Source: https://data.cityofnewyork.us/Public-Safety/NYPD-Arrest-Data-Year-to-Date-/uip8-fykc*

```{r}
df_arrest_ytd <- readr::read_csv("datasets/NYPD_Arrest_Data__Year_to_Date_.csv")
head(df_arrest_ytd)
```

## NYPD Hate Crimes

*Source: https://data.cityofnewyork.us/Public-Safety/NYPD-Hate-Crimes/bqiq-cu78*

```{r}
df_hate_crimes <- readr::read_csv("datasets/NYPD_Hate_Crimes.csv")
head(df_hate_crimes)
```

## NYC Population Data

### New York City Population By Neighborhood

*Source: https://data.cityofnewyork.us/City-Government/New-York-City-Population-By-Neighborhood-Tabulatio/swpk-hqdp*

The data was collected from Census Bureaus' Decennial data dissemination (SF1).
Neighborhood Tabulation Areas (NTAs), are aggregations of census tracts that are subsets of New York City's 55 Public Use Microdata Areas (PUMAs). Primarily due to these constraints, NTA boundaries and their associated names may not definitively represent neighborhoods.
This report shows change in population from 2000 to 2010 for each NTA.
Compiled by the Population Division – New York City Department of City Planning.

Column description:
- Standard Code Census Bureau uses to process data. The code is unique to all geographical units.
   - 005-Bronx, 047-Brooklyn, 061-Manhattan, 081-Queens, 085-Staten Island."
- Each NTA within NYC's 5 boroughs is given a unique value. 
- Neighborhood Tabulation Areas, or NTAs, are aggregations of census tracts that are subsets of New York City's 55 Public Use Microdata Areas (PUMAs).

```{r}
df_population_yearly <- readr::read_csv("datasets/New_York_City_Population_By_Neighborhood_Tabulation_Areas.csv")
head(df_population_yearly)
```

### Projected population per age group

*Source: https://data.cityofnewyork.us/City-Government/Projected-Population-2010-2040-Total-By-Age-Groups/97pn-acdf*

```{r}
df_population_age <- readr::read_csv("datasets/Projected_Population_2010-2040_-_Total_By_Age_Groups.csv")
head(df_population_age)
```

### Initial Questions

 - This census data can help identify the areas with higher crimes based on percentage of its own Population

## NYC Shooting Data 

### NYPD Shooting Incident Data (Year To Date)

*Source: https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Year-To-Date-/5ucz-vwe8*

```{r}
df_shooting_ytd <- readr::read_csv("datasets/NYPD_Shooting_Incident_Data__Year_To_Date_.csv")
head(df_shooting_ytd)
```

 - This is a breakdown of every shooting incident that occurred in NYC during the current calendar year. (2020) This data is manually extracted every quarter and reviewed by the Office of Management Analysis and Planning before being posted on the NYPD website.
Each record represents a shooting incident in NYC and includes information about the event, the location and time of occurrence. 
 - In addition, information related to suspect and victim demographics is also included. For example, has information about the sex, Age, Race of the victim.

### NYPD Shooting Incident Historic Data

*Source: https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Historic-/833y-fsy8*

```{r}
df_shooting_old <- readr::read_csv("datasets/NYPD_Shooting_Incident_Data__Historic_.csv")
head(df_shooting_old)
```
 
 
 - This is a breakdown of every shooting incident that occurred in NYC going back to 2006 through the end of the previous calendar year.(2019) This data is manually extracted every quarter and reviewed by the Office of Management Analysis and Planning before being posted on the NYPD website. Each record represents a shooting incident in NYC and includes information about the event, the location and time of occurrence. In addition, information related to suspect and victim demographics is also included.
 
### Initial Questions

 - The information about the shootings in across locations and multiple demographics can help identify groups more susceptible.
 - Further, Historical data of the shootings can help identify dangerous  neighborhoods for different groups. Possible hot-spots to avoid.
 - Also, might help analyze the overall trend in shooting across NYC over the years/months/day-of-week/time-of-day.

## FBI Human Trafficking Data

*Source: https://crime-data-explorer.fr.cloud.gov/downloads-and-docs*

```{r}
df_fbi_ht <- readr::read_csv("datasets/HT_2013_2019.csv")
head(df_fbi_ht)
```

*Many other important datasets can be found on this website.*

## FBI Drug Arrests Data

*(Reported Number of Drug Arrests)*

*Source: https://crime-data-explorer.fr.cloud.gov/downloads-and-docs*

```{r}
df_fbi_drug_arrests <- readr::read_csv("datasets/arrests_national_drug.csv")
head(df_fbi_drug_arrests)
```

## NYC citizen Census Data

*Source: https://datausa.io/profile/geo/new-york-ny#economy*

```{r}
df_nyc_earnings<- readr::read_csv("https://datausa.io/api/data?measure=Household%20Income%20by%20Race,Household%20Income%20by%20Race%20Moe&Geography=16000US3651000:parents,16000US3651000,16000US3651000:similar")
head(df_nyc_earnings)
```

## NYC park crime data

*Source: https://www1.nyc.gov/site/nypd/stats/crime-statistics/park-crime-stats.page*

```{r}
df_parks<- readr::read_csv("datasets/parks/finalpark.csv")
head(df_parks)
```