# Data transformation

## Shootings in NYC


```{r}
## Rahul
library(naniar)
library("ggplot2")
library("dplyr")
library("tidyr")
library("lubridate")
library("ggpubr")
library("data.table")
```

### Pre-processing Data


```{r, echo=FALSE}

# df_shooting_ytd <- readr::read_csv("datasets/NYPD_Shooting_Incident_Data__Year_To_Date_.csv")
# df_shooting_old <- readr::read_csv("https://data.cityofnewyork.us/resource/833y-fsy8.csv")
```


```{r, echo=TRUE}
colnames(df_shooting_old)<- toupper(colnames(df_shooting_old))
colnames(df_shooting_ytd)<- toupper(colnames(df_shooting_ytd))
```

 - Column names in the old and new data have different case, converting to a common format for merging the 2 later.

```{r}
df_shooting_ytd$OCCUR_DATE <-as.Date(df_shooting_ytd$OCCUR_DATE, format = "%m/%d/%Y")
df_shooting_old$OCCUR_DATE <-as.Date(df_shooting_old$OCCUR_DATE, format = "%m/%d/%Y")

df_shooting_total<-bind_rows(df_shooting_ytd, df_shooting_old)
df_shooting_total$OCCUR_DATE <-as.Date(df_shooting_total$OCCUR_DATE, format = "%m/%d/%Y")
```

 - Converting the date columns in both to date class using format and then merging the 2 datasets in one.
 
### Feature Engineering

 - creating more features from the existing features to help 

```{r, echo=TRUE}
# extract year and month from date - #TODO: hour and dayofweek
df_shooting_total<-df_shooting_total %>%
  mutate(year = year(OCCUR_DATE), month = lubridate::month(OCCUR_DATE,label = TRUE), day = lubridate::wday(OCCUR_DATE,label = TRUE))
```
 
  - Extracted year, month, day of the week and hour of shooting
  - #TODO!!!!!!!!!!!!!!!!!!!
  - Calculated Percentage change in number of shootings every month.

