# Missing values

```{r setup, include=FALSE}
# keep this chunk in your .Rmd file
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Human Trafficking

*For Human Trafficking data, we analyze missing values as following:*

```{r}
library(tidyverse)
library(ggplot2)
library(mi)
```

```{r}
#df_fbi_ht <- readr::read_csv("datasets/HT_2013_2019.csv")
fbiHT_df <- as.data.frame(df_fbi_ht)
x <- missing_data.frame(fbiHT_df)
image(x)
```

Alternatively, we can also view missing data as we can see below:

```{r}
colSums(is.na(df_fbi_ht))
```

## FBI Drug Arrests Data

*For FBI Drug Arrests data, we analyze missing values as following:*

```{r}
df_fbi_drug_arrests <- readr::read_csv("datasets/arrests_national_drug.csv")
drug_df <- as.data.frame(df_fbi_drug_arrests)
y <- missing_data.frame(drug_df)
image(y)
```

Alternatively, we can also view missing data as we can see below:

```{r}
colSums(is.na(df_fbi_drug_arrests))
```

## Hate Crimes 

```{r}
df_hate <- read.csv("datasets/NYPD_Hate_Crimes.csv")
df_hate$Full.Complaint.ID <- as.character(df_hate$Full.Complaint.ID)
```

We have converted complaint ID from int to character string now. 

```{r}
summary(df_hate)
```

We now explore if there are missing values in this dataset. 

```{r}
colSums(is.na(df_hate))
```

Alternate Method: 

```{r}
sapply(df_hate, function(x) sum(is.na(x)))
```

We see here that there is no missing data in this dataset. 

