# Missing values

```{r setup, include=FALSE}
# keep this chunk in your .Rmd file
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(mi)
library(naniar)
```

## Human Trafficking

*For Human Trafficking data, we analyze missing values as following:*

```{r}
#df_fbi_ht <- readr::read_csv("datasets/HT_2013_2019.csv")
fbiHT_df <- as.data.frame(df_fbi_ht)
x <- missing_data.frame(fbiHT_df)
image(x)
```

Alternatively, we can also view missing data as we can see below:

```{r}
colSums(is.na(df_fbi_ht))
```

## New York City Population

 - There are no missing values in this dataset.

## Arrests Data

```{r}
gg_miss_var(df_arrest_ytd)
```

The main feature of concern is LAW_CAT_CD which depicts level of offense (felony, misdemeanor, violation), but since there is no way to fill in this information of a categorical variable, we will ignore this feature for our analysis.

## Shootings in NYC

```{r}
gg_miss_var(df_shooting_total)
```

 - dropping the missing columns, as they cannot be imputed in anyway. We intend to focus on the victim trend in this analysis. Hence, it is fine to not consider the above columns in the current analysis.

## FBI Drug Arrests Data

*For FBI Drug Arrests data, we analyze missing values as following:*

```{r}
df_fbi_drug_arrests <- readr::read_csv("datasets/arrests_national_drug.csv")
drug_df <- as.data.frame(df_fbi_drug_arrests)
y <- missing_data.frame(drug_df)
image(y)
```

Alternatively, we can also view missing data as we can see below:

```{r}
colSums(is.na(df_fbi_drug_arrests))
```

```{r}
gg_miss_var(df_fbi_drug_arrests)
```

State_abbr has been found to be completely missing. 

## Hate Crimes 
*For Hate Crimes data, we analyze missing values as following:*
```{r}
df_hate <- read.csv("datasets/NYPD_Hate_Crimes.csv")
df_hate$Full.Complaint.ID <- as.character(df_hate$Full.Complaint.ID)
```

We have converted complaint ID from int to character string now. 

```{r}
summary(df_hate)
```

We now explore if there are missing values in this dataset. 

```{r}
colSums(is.na(df_hate))
```

Alternate Method: 

```{r}
sapply(df_hate, function(x) sum(is.na(x)))
mdhate <- missing_data.frame(df_hate)
image(mdhate)
```


```{r}
gg_miss_var(mdhate)
```

Arrest.Id and Arrest.Date have the same missing pattern. Other.Motive.Description has been found to have more than 700 missing values.

## Park Crime

*For Park Crime data, we analyze missing values as following:*

```{r}
df_parks <- read.csv('datasets/parks/finalpark.csv')
colSums(is.na(df_parks))
```


```{r}
mdparks <- missing_data.frame(df_parks)
gg_miss_var(mdparks)
```

None of the variables in this dataset are missing. 