# Missing values

```{r setup, include=FALSE}
# keep this chunk in your .Rmd file
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<!-- ```{r setup, include=FALSE} -->
<!-- ## Rahul setup -->
<!-- df_shooting_ytd <- readr::read_csv("datasets/NYPD_Shooting_Incident_Data__Year_To_Date_.csv") -->
<!-- df_shooting_old <- readr::read_csv("https://data.cityofnewyork.us/resource/833y-fsy8.csv") -->
<!-- ``` -->

## Human Trafficking

*For Human Trafficking data, we analyze missing values as following:*

```{r}
library(tidyverse)
library(ggplot2)
library(mi)
```

```{r}
#df_fbi_ht <- readr::read_csv("datasets/HT_2013_2019.csv")
fbiHT_df <- as.data.frame(df_fbi_ht)
x <- missing_data.frame(fbiHT_df)
image(x)
```

Alternatively, we can also view missing data as we can see below:

```{r}
colSums(is.na(df_fbi_ht))
```

## New York City Population

 - There are no missing values in this dataset.

## Shootings in NYC

```{r}
gg_miss_var(df_shooting_total)
```

 - dropping the missing columns, as they cannot be imputed in anyway. We intend to focus on the victim trend in this analysis. Hence, it is fine to not consider the above columns in the current analysis.

## FBI Drug Arrests Data

*For FBI Drug Arrests data, we analyze missing values as following:*

```{r}
df_fbi_drug_arrests <- readr::read_csv("datasets/arrests_national_drug.csv")
drug_df <- as.data.frame(df_fbi_drug_arrests)
y <- missing_data.frame(drug_df)
image(y)
```

Alternatively, we can also view missing data as we can see below:

```{r}
colSums(is.na(df_fbi_drug_arrests))
```


## Hate Crimes 

```{r}
df_hate <- read.csv("datasets/NYPD_Hate_Crimes.csv")
df_hate$Full.Complaint.ID <- as.character(df_hate$Full.Complaint.ID)
```

We have converted complaint ID from int to character string now. 

```{r}
summary(df_hate)
```

We now explore if there are missing values in this dataset. 

```{r}
colSums(is.na(df_hate))
```

Alternate Method: 

```{r}
sapply(df_hate, function(x) sum(is.na(x)))
```

We see here that there is no missing data in this dataset. 

