# Results

## Shootings in NYC

### Trend in number of shootings

#### yearly 
#### monthly or yearly (DISCUSS)

##### Whole NYC

```{r}
df_shoot_grouped2 <- df_shooting_total %>%
  select(year, month, STATISTICAL_MURDER_FLAG, VIC_AGE_GROUP, VIC_SEX, VIC_RACE) %>%
  group_by(year, month, STATISTICAL_MURDER_FLAG, VIC_SEX, VIC_AGE_GROUP, VIC_RACE) %>%
  summarize(occurances = n())

df_shoot_grouped2$date<-as.Date(paste(df_shoot_grouped2$year, df_shoot_grouped2$month,"1",sep="-"), format = "%Y-%b-%d")

temp_df<- df_shoot_grouped2 %>% 
  filter(year>=2019) %>%
  group_by(date, month) %>%
  summarize(total_shots= sum(occurances))#%>%
  # mutate(monthly_change = (total_shots/shift(total_shots, n=1, fill = NA, type="lag") -1)*100)

temp_df$monthly_change<- ((temp_df$total_shots/shift(temp_df$total_shots, n=1, fill=NA, type="lag"))-1)
```

```{r, echo=TRUE, fig.width=10,fig.height=11}
ggplot(temp_df, aes(x= date)) + 
  geom_bar(aes(y = monthly_change, fill = factor(month)), stat="identity") +
  geom_text(aes( label = scales::percent(monthly_change),
                  y= monthly_change ), stat= "identity",angle = 90) +
  scale_y_continuous(labels = scales::percent) +
  ggtitle("Is it that time of the year?!",
          subtitle = "Percentage change in number of Shootings by month") +
  labs(y = "Percentage Change", x="Date", fill="Month", caption = "Source: NYC-open-Data::Shooting Incident Data") +
  theme(plot.title = element_text(face = "bold")) +
  theme(plot.subtitle = element_text(face = "bold", color = "grey35")) +
  theme(plot.caption = element_text(color = "grey68"))
```

 - We see every year in the month of May and June the number of shootings rise.
 - Last year, (possibly due to covid restrictions easing up) saw 92-95% increase in these months.

#### weekly

```{r}
df_shoot_grouped_week <- df_shooting_total %>%
  select(OCCUR_DATE, day, STATISTICAL_MURDER_FLAG, VIC_AGE_GROUP, VIC_SEX, VIC_RACE) %>%
  group_by(OCCUR_DATE, day, STATISTICAL_MURDER_FLAG, VIC_SEX, VIC_AGE_GROUP, VIC_RACE) %>%
  summarize(occurances = n())

# df_shoot_grouped_week$date<-as.Date(paste(df_shoot_grouped_Week$year, df_shoot_grouped_Week$month,"1",sep="-"), format = "%Y-%b-%d")

temp_df<- df_shoot_grouped_week %>% 
  group_by(day) %>%
  summarize(total_shots= mean(occurances))#%>%
  # mutate(monthly_change = (total_shots/shift(total_shots, n=1, fill = NA, type="lag") -1)*100)

temp_df$weekly_change<- ((temp_df$total_shots/shift(temp_df$total_shots, n=1, fill=NA, type="lag"))-1)
```

```{r, echo=TRUE}
ggplot(temp_df, aes(x= day, y = total_shots, fill = factor(day), label=sprintf("%0.2f", round(total_shots, digits = 2)))) + 
  geom_bar(stat="identity") +
  geom_text( vjust=-.5) +
  ggtitle("Does the day of the week matter?!",
          subtitle = "Average number of shootings on the day of the week") +
  labs(y = "Average shooting count", x="Day", fill="Weekday", caption = "Source: NYC-open-Data::Shooting Incident Data") +
  theme(plot.title = element_text(face = "bold")) +
  theme(plot.subtitle = element_text(face = "bold", color = "grey35")) +
  theme(plot.caption = element_text(color = "grey68"))
```
#### daily

### Trend in number of deaths

### Susciptible groups

### Hot-spot neighbourhoods/AREAS 

 - (discuss!)

```{r}
library(tidyverse)
library(ggplot2)
library(mi)

#df_fbi_ht <- readr::read_csv("datasets/HT_2013_2019.csv")

# fbiHT_df <- as.data.frame(df_fbi_ht)

# summary(df_fbi_ht)

theme_dotplot <- theme_bw(14) +
  theme(axis.text.y = element_text(size = rel(.75)),
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(size = rel(.75)),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(size = 0.5),
        panel.grid.minor.x = element_blank())
```

## Human Trafficking

```{r}
df_fbi_ht_temp <- df_fbi_ht %>%
  group_by(STATE_NAME) %>%
  summarise(Total = n())

ggplot(df_fbi_ht_temp, aes(x = Total, y = reorder(STATE_NAME, Total))) +
  geom_point(color = "blue") +
  theme_dotplot +
  xlab("Total registered cases") +
  ylab("State") +
  ggtitle("Registered Human Trafficking cases by State (2013-19)")
```

## Hate Crimes 

```{r}
hate3 <- ggplot(data=df_hate, aes(x=County, fill=factor(Complaint.Year.Number))) + geom_bar(data=df_hate)
hate3
```

#### Insights
Kings County (Brooklyn) had the highest total number of hate crimes over the years 2019 and 2020. This was followed by New York (Manhattan) and then Queens. 

The only county/borough to have experienced an upturn in hate crimes in 2020  was the Richmond county (Staten Island Borough). 

This is confirmed by articles such as (Source: silive.com https://www.silive.com/crime-safety/2021/04/staten-island-was-the-only-borough-in-nyc-with-an-increase-in-hate-crimes-in-2020-why-where-were-they-crime-analysis.html)


```{r}
library(ggplot2)
df_hate <- read.csv("datasets/NYPD_Hate_Crimes.csv")
hate1 <- ggplot(data=df_hate, aes(y=Offense.Category, fill=County)) + geom_bar(data=df_hate, alpha=0.9, position = position_stack(reverse = TRUE))
hate1
```

#### Insights 
The largest proportion of hate crimes stem from religious reasons, followed by race and sexual orientation. 

Kings County (Brooklyn) has a much larger proportion of these religious crimes. Interestingly, although the total number of hate crimes is greater in Kings County (Brooklyn) than New York County (Manhattan), the number of crimes stemming from Sexual Orientation and Race/Color is greater in New York County than Kings County. 

```{r}
theme_dotplot <- theme_bw(14) +
    theme(axis.text.y = element_text(size = rel(.5)),
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(size = rel(.75)),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(size = 1),
        panel.grid.minor.x = element_blank())

df_temp <- as.data.frame(table(df_hate$Bias.Motive.Description))
#View(df_temp)
hate2 <- ggplot(df_temp, aes(x = Freq, 
               y = reorder(Var1, Freq))) +
    geom_point(color = "blue") +
    theme_dotplot +
    xlab("Incident Count") +
    ylab("Motive Type") +
    ggtitle("Motive Type Counts")
hate2
```

#### Insights

Anti- Jewish hate crimes seem to be the largest in number (by a large amount, almost 350 crimes) during this time period, followed by anti-homosexual (gay) and anti-black crimes. 

```{r}
hate4 <- ggplot(data=df_hate, aes(y=Bias.Motive.Description, fill=County)) + geom_bar(data=df_hate, alpha=0.9, position = position_stack(reverse = TRUE))
hate4
```

#### Insights

As expected from the previous 2 graphs, Kings County (Brooklyn) has the highest number of anti-jewish crimes. Kings County has a disproportionately large number of these crimes, whereas other motive types like anti-black or anti-homosexual seem to be equal in the boroughs of Kings and New York (Manhattan).

These insights are confirmed by articles such as (https://www.cbsnews.com/news/nypd-to-increase-presence-in-brooklyn-after-string-of-anti-semitic-hate-crimes/ Source: CBS News)


```{r}
df_hate$Complaint.Year.Number <- as.factor(df_hate$Complaint.Year.Number)
hate5 <- ggplot(data=df_hate, aes(x=Complaint.Year.Number, fill=Offense.Category)) + geom_bar(data=df_hate)
hate5
```

#### Insights
Across the city of New York, the number of hate crimes went down in 2020 (by almost 40%). Sexual orientation crimes decreased, whereas crimes based on race/color unfortunately went up in number. 

```{r}
df_temp <- as.data.frame(table(df_hate$Month.Number, df_hate$Complaint.Year.Number))

df_temp$new_col <- c(12:1)
#View(df_temp)
hate6 <- ggplot(df_temp, aes(x = Freq, 
               y = reorder(Var1, new_col), color=Var2)) +
    geom_point() +
    theme_dotplot +
    xlab("Incident Count") +
    ylab("Month Number") +
    ggtitle("Incident Counts by month")
hate6
```

#### Insights
Across both years, it is visible that there is a dependance on the month when it comes to the number of crimes. 

However, for all months, 2020 had lower number of hate crimes than there were in 2019. The highest crime number for both years was in the month of March. 

```{r}
library(vcd)
hate7 <- vcd::doubledecker(Bias.Motive.Description ~ Complaint.Year.Number + County, data=df_hate, main='Analysing Motive with year and county') 
```

#### Insights

The insights from this mosaic plot corroborate our findings from the previous plots. 

```{r}
df_hate$date_diff <- as.Date(as.character(df_hate$Arrest.Date), format="%m/%d/%Y")-
                  as.Date(as.character(df_hate$Record.Create.Date), format="%m/%d/%Y")
hate8 <- ggplot(data=df_hate, mapping=aes(x=date_diff)) + geom_histogram( bins=20, color="#e9ecef", alpha=0.6, position = 'identity') + facet_wrap(~Complaint.Year.Number) #+ xlim(0,200)
hate8
```

#### Insights
The majority of arrests, if made, were made within 50 days since the report creation, for both the years 2019 and 2020. This indicates a heavy right-skew in the graphs. 

Only very few arrests were made beyond the 50-day mark in both years.

```{r}
hate9 <- ggplot(data=df_hate, aes(y=PD.Code.Description, fill=County)) + geom_bar(data=df_hate, alpha=0.9, position = position_stack(reverse = TRUE))
hate9
```


#### Insights
Criminal Mischief, Aggravated Assault and Aggravated Harrassment were the most code of hate crimes committed.


```{r}
x <- data.frame(table(df_hate$Arrest.Date))
length(x$Freq)
arr <- sum(x$Freq[c(2:length(x$Freq))])
narr <- x$Freq[1]
arrvec <- c(arr, narr)
labels1 <- c('Arrested', 'Not Arrested')
df <- data.frame(labs = labels1, outcome = arrvec)
ggplot() + geom_col(df, mapping = aes(labs, outcome))
```

#### Insights
Unfortunately, the majority of offenders in hate crimes during 2019 and 2020 have not been arrested. 

506 offenders were not arrested, whereas 222 arrests were made. 

This picture might be offset in the case of one person commited multiple hate crimes and then being arrested. As we do not have information regarding arrest details, such analysis is omitted here. 

## Drugs

```{r}
df_drugs <- readr::read_csv("datasets/arrests_national_drug.csv")
df_drugs <- as.data.frame(df_drugs)
View(df_drugs)
```

### Arrests related to all drugs throughout US 
```{r}
library(ggplot2)
drug1 <- ggplot()+
  geom_line(data=df_drugs, aes(x = year, y = total_arrests, colour='blue')) +
  ggtitle("Arrests related to drugs across US across year") +
  labs(x = "year", y = "number of arrests") + scale_color_discrete(name = "Arrest Count", labels = c("total arrests"))
drug1
```
### Arrests related to manufacture and possession
```{r}
drug2 <- ggplot()+
  geom_line(data=df_drugs, aes(x = year, y = total_manufacture, colour='blue')) +
  geom_line(data=df_drugs, aes(x = year, y = total_possess, colour='red'))+
  ggtitle("Arrests related to drug manufacture & possessions- US YoY") +
  labs(x = "year", y = "number of arrests") + scale_color_discrete(name = "Arrest Count", labels = c("Manufacture", "Possession"))
drug2
```


### Arrests related to Manufacturing 
```{r}
drug3 <- ggplot()+
  geom_line(data=df_drugs, aes(x = year, y = opioid_manufacture, colour='blue')) +
  geom_line(data=df_drugs, aes(x = year, y = marijuana_manufacture, colour='red'))+
  geom_line(data=df_drugs, aes(x = year, y = synthetic_manufacture, colour='orange'))+
  geom_line(data=df_drugs, aes(x = year, y = other_manufacture, colour='green'))+
  ggtitle("Arrests related to drug manufacture -US YoY") +
  labs(x = "year", y = "number of arrests") + scale_color_discrete(name = "Arrest Count", labels = c("Opioid Manufacture", "Marijuana Manufacture", 'Synthetic Manufacture', 'Other Manufacture'))
drug3
```

### Arrests related to Possession

```{r}
drug4 <- ggplot()+
  geom_line(data=df_drugs, aes(x = year, y = opioid_possess, colour='blue')) +
  geom_line(data=df_drugs, aes(x = year, y = marijuana_possess, colour='red'))+
  geom_line(data=df_drugs, aes(x = year, y = synthetic_possess, colour='orange'))+
  geom_line(data=df_drugs, aes(x = year, y = other_possess, colour='green'))+
  ggtitle("Arrests related to drug manufacture -US YoY") +
  labs(x = "year", y = "number of arrests") + scale_color_discrete(name = "Arrest Count", labels = c("Opioid Possession", "Marijuana Possession", 'Synthetic Possession', 'Other Possession'))
drug4
```



